---
output: html_document
---

<style>
.title-slide {
  height: 100vh;
  display: flex;
  flex-direction: column;
  justify-content: center;
  text-align: center;
}

.title-slide h1 {
  font-size: 42px;
}

.title-slide h2 {
  font-size: 24px;
  font-weight: normal;
}

.title-slide p {
  font-size: 20px;
}
</style>

<div class="title-slide">

# World Happiness Report Analysis (2017–2020)

## Data Visualization for Analytics

### Anna Reschko

Professor: Dr. Shiju Zhang 

STAT 660 

May 2026

</div>

\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


## Introduction

The World Happiness Report is an annual publication that ranks countries
by self-reported life satisfaction. This project explores happiness data
from 2017 to 2020, scraped directly from Wikipedia, to answer 8 research
questions about global well-being, regional patterns, and the role of
factors like GDP, social support, and life expectancy.

## Data

Data was scraped from the Wikipedia World Happiness Report page using
the `rvest` package in R. Four years of data (2017–2020) were extracted,
cleaned, and combined into a single dataset of 620 country-year observations.

```{r packages}

# Load Libraries
library(rvest)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sqldf)
library(treemap)
library(maps)
library(ggrepel)
library(RColorBrewer)
library(kableExtra)
```


```{r data}

# Data Collection and Preparation 
# This section contains the setup and preprocessing code used to build the
# combined World Happiness dataset for the analysis.

# The code:
# 1. Pulls World Happiness Report tables from Wikipedia
# 2. Extracts relevant variables for each year (2017–2020)
# 3. Standardizes country names across datasets
# 4. Combines all yearly datasets into one master dataset


tabs <- read_html("https://en.wikipedia.org/wiki/World_Happiness_Report") |>
  html_table(fill = TRUE)

get_year <- function(tab, rank_col, country_col, year) {
  tab |>
    select(Rank = all_of(rank_col), Country = all_of(country_col),
           Score, GDP = `GDP per capita`,
           Social = `Social support`,
           Life = `Healthy life expectancy`) |>
    mutate(across(c(Rank, Score, GDP, Social, Life), as.numeric),
           Year = year, Country = trimws(Country)) |>
    filter(!is.na(Score), !is.na(Rank))
}

fix_names <- function(df) {
  df |> mutate(Country = recode(Country,
    "United States of America" = "United States",
    "Russian Federation"       = "Russia",
    "Trinidad & Tobago"        = "Trinidad and Tobago",
    "North Macedonia"          = "Macedonia",
    "Czech Republic"           = "Czechia"))
}

h2020 <- fix_names(get_year(tabs[[23]], "Overall rank", "Country or region", 2020))
h2019 <- fix_names(get_year(tabs[[25]], "Overall rank", "Country or region", 2019))
h2018 <- fix_names(get_year(tabs[[27]], "Overall rank", "Country or region", 2018))
h2017 <- fix_names(get_year(tabs[[29]], "Overall Rank", "Country",           2017))

all_years <- bind_rows(h2017, h2018, h2019, h2020)

# Preview datasets

knitr::kable(head(h2017),
             caption = "Preview of 2017 Happiness Dataset")
knitr::kable(head(h2018),
             caption = "Preview of 2018 Happiness Dataset")

knitr::kable(head(h2019),
             caption = "Preview of 2019 Happiness Dataset")

knitr::kable(head(h2020),
             caption = "Preview of 2020 Happiness Dataset")

knitr::kable(head(all_years),
             caption = "Preview of Combined Dataset")

```

## Results

### Question 1 - Top 10 Happiest Countries in 2020

Finland ranks as the happiest country in 2020 with a score of 7.81,
followed closely by Denmark and Switzerland. All top 10 countries are
high-income nations, with 7 of the 10 being Nordic or Western European.

```{r q1}

# Question 1: Top 10 happiest countries in 2020 

h2020 |>
  arrange(desc(Score)) |>
  slice(1:10) |>
  ggplot(aes(x = reorder(Country, -Score), y = Score, fill = Score)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  scale_fill_gradient(low = "#c7f0c2", high = "#006400") +
  geom_text(aes(label = round(Score, 2)), vjust = -0.4, size = 3.2) +
  labs(title = "Top 10 Happiest Countries (2020)",
       x = NULL, y = "Happiness Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

### Question 2 - Happiness Trends for 6 Major Countries (2017–2020): China, India, USA, Indonesia, Japan, and Russia

The United States shows a slight declining trend across all 4 years.
China and Russia remain relatively stable. India consistently scores
the lowest of the 6 countries, while Japan sits in the mid-range
despite its high GDP.

```{r q2}

# Question 2: Happiness score trends for selected countries (2017–2020) 

all_years |>
  filter(Country %in% c("China", "India", "United States",
                         "Indonesia", "Japan", "Russia")) |>
  ggplot(aes(x = Year, y = Score, color = Country, group = Country)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = 2017:2020) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Trends in World Happiness Scores (2017–2020)",
    subtitle = "Selected countries from the World Happiness Report",
    x = "Year",
    y = "Happiness Score",
    color = "Country"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```


### Question 3 - Consistently Top 10 Countries

The same core group of Nordic countries — Finland, Denmark, Norway,
Iceland, Switzerland, Netherlands, and Sweden — appear in the top 10
in all 4 years, suggesting stable structural advantages.

```{r q3}

# Question 3: Countries consistently in the top 10 (2017–2020) 

consistent <- all_years |>
  group_by(Year) |>
  slice_max(Score, n = 10) |>
  ungroup() |>
  count(Country, name = "Years_in_Top10") |>
  filter(Years_in_Top10 == 4)

knitr::kable(
  consistent,
  caption = "Countries Ranked in the Top 10 Across All Four Years"
)

consistent |>
  ggplot(aes(x = reorder(Country, Years_in_Top10),
             y = Years_in_Top10)) +
  geom_bar(stat = "identity",
           fill = "darkgreen",
           width = 0.7) +
  coord_flip() +
  scale_y_continuous(breaks = 0:4, limits = c(0,4)) +
  labs(
    title = "Countries Ranked in the Top 10 Across All Years (2017–2020)",
    subtitle = "Based on World Happiness Report scores",
    x = NULL,
    y = "Number of Years in Top 10"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

### Question 4 - GDP vs Happiness Score Correlation

There is a strong positive correlation between GDP per capita and
happiness score across all 4 years. The relationship is consistent
year to year, indicating GDP is a reliable but not exclusive predictor
of national happiness.

```{r q4}

# Question 4: GDP per capita vs Happiness score by Year (sqldf)

sqldf("SELECT Country, Score, GDP, Year
       FROM all_years
       WHERE GDP IS NOT NULL AND Score IS NOT NULL") |>
  ggplot(aes(x = GDP, y = Score, color = factor(Year))) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(method = "lm",
              se = FALSE,
              aes(group = factor(Year)),
              linewidth = 1.1) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Relationship Between GDP per Capita and Happiness Scores",
    subtitle = "World Happiness Report data (2017–2020)",
    x = "Log GDP per Capita",
    y = "Happiness Score",
    color = "Year"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# Correlation Table between GDP and Happiness by Year 

correlation_table <- sqldf("
    SELECT Country, Score, GDP, Year
    FROM all_years
    WHERE GDP IS NOT NULL AND Score IS NOT NULL
  ") |>
  group_by(Year) |>
  summarise(
    Correlation = round(cor(GDP, Score, use = "complete.obs"), 3)
  )

knitr::kable(
  correlation_table,
  caption = "Correlation Between GDP per Capita and Happiness Score by Year"
)
```


### Question 5 - Regional Distribution of Happiness (2020)

Europe has the highest median happiness score and least variability.
Africa has the lowest median with wide spread. The Americas show
moderate scores with notable outliers.

```{r q5}

# Question 5: Distribution of happiness scores by region 2020 

data("GNI2014", package = "treemap")

h2020 |>
  left_join(GNI2014 |> select(Country = country, Region = continent),
            by = "Country") |>
  mutate(Region = case_when(
    Country == "United States"                    ~ "North America",
    Country %in% c("Russia", "Kosovo",
                   "North Cyprus")                ~ "Europe",
    Country %in% c("Taiwan", "Hong Kong")         ~ "Asia",
    is.na(Region)                                 ~ "Other",
    TRUE                                          ~ Region
  )) |>
  filter(Region != "Other") |>
  ggplot(aes(x = reorder(Region, Score, FUN = median),
             y = Score, fill = Region)) +
  geom_boxplot(show.legend = FALSE, outlier.colour = "tomato") +
  labs(title = "Happiness Score Distribution by Region (2020)",
       x = NULL, y = "Happiness Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

### Question 6 - Social Support vs Life Expectancy Impact on Happiness (2020)

Countries with both high life expectancy and high social support
tend to score highest on happiness. High life expectancy alone
does not guarantee high happiness without strong social support.

```{r q6}
# Question 6: Bubble plot - social support vs life expectancy 2020 

ggplot(h2020, aes(x = Life, y = Social, size = Score, color = Score)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(2, 14), name = "Happiness Score") +
  scale_color_gradient(low = "#c7f0c2", high = "#006400", name = "Happiness Score") +
  labs(title = "Social Support and Life Expectancy in Relation to Happiness",
    subtitle = "World Happiness Report data (2020)",
    x = "Healthy Life Expectancy",
    y = "Social Support") +
  theme_minimal()


```

### Question 7 - Treemap (2020)

```{r q7}

# Question 7: Treemap for 2020 happiness rankings 

h2020 |>
  arrange(desc(Score)) |>
  mutate(Rank = row_number()) |>

  treemap(
    index             = "Country",
    vSize             = "Score",
    vColor            = "Rank",
    type              = "value",
    palette           = rev(brewer.pal(9, "RdYlGn")),

    title             = "Treemap of World Happiness Rankings (2020)",
    title.legend      = "Happiness Rank\n(1 = Happiest)",

    fontsize.labels   = 10,
    fontcolor.labels  = "black",
    inflate.labels    = FALSE
  )
```

### Question 8 - Choropleth Map (2020)

The map clearly shows the geographic happiness divide — Northern
Europe dominates in green while much of Africa and parts of Asia
appear in red. Gray areas indicate missing data.


```{r q8}
# Question 8: Choropleth map 2020 

h2020 |>
  mutate(Country = recode(Country,
    "United States"        = "USA",
    "United Kingdom"       = "UK",
    "Trinidad and Tobago"  = "Trinidad",
    "North Cyprus"         = "Cyprus",
    "Congo"                = "Republic of Congo",
    "DR Congo"             = "Democratic Republic of the Congo"
  )) |>
  merge(map_data("world"), by.x = "Country", by.y = "region", all.y = TRUE) |>
  arrange(group, order) |>
  ggplot(aes(x = long, y = lat, group = group, fill = Score)) +
  geom_polygon(color = "white", linewidth = 0.1) +
  coord_fixed(ratio = 1.25) +
  scale_fill_gradient(low = "#e41a1c",
high = "#006400",
                      na.value = "gray80", name = "Happiness\nScore") +
  labs(title = "Global Distribution of Happiness Scores (2020)") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Conclusion

This analysis confirms that happiness is strongly associated with
GDP, social support, and life expectancy. Nordic countries consistently
lead global rankings. Regional disparities are large and persistent,
with Europe far outpacing Africa across all metrics. The 2017–2020
period shows relative stability in rankings with modest year-to-year
changes for most countries.

## Appendix

All code used in this project was written in R and is visible in the
chunks below. Data was sourced from:

Wikipedia - World Happiness Report:
https://en.wikipedia.org/wiki/World_Happiness_Report


### Code for Project Setup and Data Preparation:

```{r appendix-packages, echo=TRUE, eval=FALSE}

# Load Libraries
library(rvest)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sqldf)
library(treemap)
library(maps)
library(ggrepel)
library(RColorBrewer)
library(kableExtra)
```

```{r appendix-data, echo=TRUE, eval=FALSE}

# Data Collection and Preparation 
# This section contains the setup and preprocessing code used to build the
# combined World Happiness dataset for the analysis.

# The code:
# 1. Pulls World Happiness Report tables from Wikipedia
# 2. Extracts relevant variables for each year (2017–2020)
# 3. Standardizes country names across datasets
# 4. Combines all yearly datasets into one master dataset


tabs <- read_html("https://en.wikipedia.org/wiki/World_Happiness_Report") |>
  html_table(fill = TRUE)

get_year <- function(tab, rank_col, country_col, year) {
  tab |>
    select(Rank = all_of(rank_col), Country = all_of(country_col),
           Score, GDP = `GDP per capita`,
           Social = `Social support`,
           Life = `Healthy life expectancy`) |>
    mutate(across(c(Rank, Score, GDP, Social, Life), as.numeric),
           Year = year, Country = trimws(Country)) |>
    filter(!is.na(Score), !is.na(Rank))
}

fix_names <- function(df) {
  df |> mutate(Country = recode(Country,
    "United States of America" = "United States",
    "Russian Federation"       = "Russia",
    "Trinidad & Tobago"        = "Trinidad and Tobago",
    "North Macedonia"          = "Macedonia",
    "Czech Republic"           = "Czechia"))
}

h2020 <- fix_names(get_year(tabs[[23]], "Overall rank", "Country or region", 2020))
h2019 <- fix_names(get_year(tabs[[25]], "Overall rank", "Country or region", 2019))
h2018 <- fix_names(get_year(tabs[[27]], "Overall rank", "Country or region", 2018))
h2017 <- fix_names(get_year(tabs[[29]], "Overall Rank", "Country",           2017))

all_years <- bind_rows(h2017, h2018, h2019, h2020)

# Preview datasets

knitr::kable(head(h2017),
             caption = "Preview of 2017 Happiness Dataset")
knitr::kable(head(h2018),
             caption = "Preview of 2018 Happiness Dataset")

knitr::kable(head(h2019),
             caption = "Preview of 2019 Happiness Dataset")

knitr::kable(head(h2020),
             caption = "Preview of 2020 Happiness Dataset")

knitr::kable(head(all_years),
             caption = "Preview of Combined Dataset")


```

### Code for Question Results:

```{r appendix-q1, echo=TRUE, eval=FALSE}

# Question 1: Top 10 happiest countries in 2020 

h2020 |>
  arrange(desc(Score)) |>
  slice(1:10) |>
  ggplot(aes(x = reorder(Country, -Score), y = Score, fill = Score)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  scale_fill_gradient(low = "#c7f0c2", high = "#006400") +
  geom_text(aes(label = round(Score, 2)), vjust = -0.4, size = 3.2) +
  labs(title = "Top 10 Happiest Countries (2020)",
       x = NULL, y = "Happiness Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

```

```{r appendix-q2, echo=TRUE, eval=FALSE}

# Question 2: Happiness score trends for selected countries (2017–2020) 

all_years |>
  filter(Country %in% c("China", "India", "United States",
                         "Indonesia", "Japan", "Russia")) |>
  ggplot(aes(x = Year, y = Score, color = Country, group = Country)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = 2017:2020) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Trends in World Happiness Scores (2017–2020)",
    subtitle = "Selected countries from the World Happiness Report",
    x = "Year",
    y = "Happiness Score",
    color = "Country"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```

```{r appendix-q3, echo=TRUE, eval=FALSE}

# Question 3: Countries consistently in the top 10 (2017–2020) 

consistent <- all_years |>
  group_by(Year) |>
  slice_max(Score, n = 10) |>
  ungroup() |>
  count(Country, name = "Years_in_Top10") |>
  filter(Years_in_Top10 == 4)

knitr::kable(
  consistent,
  caption = "Countries Ranked in the Top 10 Across All Four Years"
)

consistent |>
  ggplot(aes(x = reorder(Country, Years_in_Top10),
             y = Years_in_Top10)) +
  geom_bar(stat = "identity",
           fill = "darkgreen",
           width = 0.7) +
  coord_flip() +
  scale_y_continuous(breaks = 0:4, limits = c(0,4)) +
  labs(
    title = "Countries Ranked in the Top 10 Across All Years (2017–2020)",
    subtitle = "Based on World Happiness Report scores",
    x = NULL,
    y = "Number of Years in Top 10"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

```{r appendix-q4, echo=TRUE, eval=FALSE}

# Question 4: GDP per capita vs Happiness score by Year (sqldf)

sqldf("SELECT Country, Score, GDP, Year
       FROM all_years
       WHERE GDP IS NOT NULL AND Score IS NOT NULL") |>
  ggplot(aes(x = GDP, y = Score, color = factor(Year))) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(method = "lm",
              se = FALSE,
              aes(group = factor(Year)),
              linewidth = 1.1) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Relationship Between GDP per Capita and Happiness Scores",
    subtitle = "World Happiness Report data (2017–2020)",
    x = "Log GDP per Capita",
    y = "Happiness Score",
    color = "Year"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# Correlation Table between GDP and Happiness by Year 

correlation_table <- sqldf("
    SELECT Country, Score, GDP, Year
    FROM all_years
    WHERE GDP IS NOT NULL AND Score IS NOT NULL
  ") |>
  group_by(Year) |>
  summarise(
    Correlation = round(cor(GDP, Score, use = "complete.obs"), 3)
  )

knitr::kable(
  correlation_table,
  caption = "Correlation Between GDP per Capita and Happiness Score by Year"
)
```

```{r appendix-q5, echo=TRUE, eval=FALSE}

# Question 5: Distribution of happiness scores by region 2020 

data("GNI2014", package = "treemap")

h2020 |>
  left_join(GNI2014 |> select(Country = country, Region = continent),
            by = "Country") |>
  mutate(Region = case_when(
    Country == "United States"                    ~ "North America",
    Country %in% c("Russia", "Kosovo",
                   "North Cyprus")                ~ "Europe",
    Country %in% c("Taiwan", "Hong Kong")         ~ "Asia",
    is.na(Region)                                 ~ "Other",
    TRUE                                          ~ Region
  )) |>
  filter(Region != "Other") |>
  ggplot(aes(x = reorder(Region, Score, FUN = median),
             y = Score, fill = Region)) +
  geom_boxplot(show.legend = FALSE, outlier.colour = "tomato") +
  labs(title = "Happiness Score Distribution by Region (2020)",
       x = NULL, y = "Happiness Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r appendix-q6, echo=TRUE, eval=FALSE}

# Question 6: Bubble plot - social support vs life expectancy 2020 

ggplot(h2020, aes(x = Life, y = Social, size = Score, color = Score)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(2, 14), name = "Happiness Score") +
  scale_color_gradient(low = "#c7f0c2", high = "#006400", name = "Happiness Score") +
  labs(title = "Social Support and Life Expectancy in Relation to Happiness",
    subtitle = "World Happiness Report data (2020)",
    x = "Healthy Life Expectancy",
    y = "Social Support") +
  theme_minimal()
```

```{r appendix-q7, echo=TRUE, eval=FALSE}

# Question 7: Treemap for 2020 happiness rankings 

h2020 |>
  arrange(desc(Score)) |>
  mutate(Rank = row_number()) |>

  treemap(
    index             = "Country",
    vSize             = "Score",
    vColor            = "Rank",
    type              = "value",
    palette           = rev(brewer.pal(9, "RdYlGn")),

    title             = "Treemap of World Happiness Rankings (2020)",
    title.legend      = "Happiness Rank\n(1 = Happiest)",

    fontsize.labels   = 10,
    fontcolor.labels  = "black",
    inflate.labels    = FALSE
  )
```

```{r appendix-q8, echo=TRUE, eval=FALSE}

# Question 8: Choropleth map 2020 

h2020 |>
  mutate(Country = recode(Country,
    "United States"        = "USA",
    "United Kingdom"       = "UK",
    "Trinidad and Tobago"  = "Trinidad",
    "North Cyprus"         = "Cyprus",
    "Congo"                = "Republic of Congo",
    "DR Congo"             = "Democratic Republic of the Congo"
  )) |>
  merge(map_data("world"), by.x = "Country", by.y = "region", all.y = TRUE) |>
  arrange(group, order) |>
  ggplot(aes(x = long, y = lat, group = group, fill = Score)) +
  geom_polygon(color = "white", linewidth = 0.1) +
  coord_fixed(ratio = 1.25) +
  scale_fill_gradient(low = "#e41a1c",
high = "#006400",
                      na.value = "gray80", name = "Happiness\nScore") +
  labs(title = "Global Distribution of Happiness Scores (2020)") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```











